### 2018-07-10 source/Ch6-loresEGS/lowres-EGS.tex
backup

\chapter{The Groth Strip with Dutch LOFAR}\label{section.EGS.lowres}
\minitoc
% % % % % % % % % % % % % % % % % % % % % % %
\section{Aims \& Methodology}

@@

% % % % % % % % % % % % % % % % % % % % % % %
\section{Data Reduction}

\subsection{Data \& Observation Properties}
\pg
The dataset used for this PhD project is part of the LOFAR Surveys KSP Tier-1 survey, which consists of a number of 8-hour pointings covering as much of the sky visibile to LOFAR as possible. We analyse one of these pointings, an observation performed on the 28th of August 2014 and centred on the Extended Groth Strip. As part of the Tier-1 survey, it is an 8-hour-long observation. We limit ourselves to analysing only the HBA observation, meaning that we use 365 sub-bands which sample a total bandwidth of 110-182 MHz. We use all Core and Remote LOFAR stations, as well as some of the International stations which were online at the time (specifically the German stations 1-5 and 7, along with the Swedish and the British station).

@@

\pg
The data was acquired through the LOFAR Long-Term Archive, and is thus pre-processed and flagged for RFI using the standard tools (NDPPP\citetads{2018ascl.soft04003V} and AOFlagger \citetads{2010ascl.soft10017O}, respectively). The next step is therefore to calibrate the data.

@@

\pg
What's more, antenna gains can vary over short durations or show peculiar spectral behaviour. Even if interferometric data were averaged massively using baseline-dependent window functions, information on this underlying gain behaviour would be lost - this means that averaging data to reduce its size (and associated reduction wall time) is a risky proposition.

@@

\pg
The calibration strategy consists of a first round of direction-independent calibration followed by direction-dependent self-calibration. In effect, we use the following RIME:
\begin{equation}
\Vis_{pq} = \Gjones_p \bigg( \sum_s \Ejones_{sp} \Kjones_{sp} \Bmatrix_{s} \Kjones_{sq}^H \Ejones_{sq}^H \bigg) \Gjones_q^H
\end{equation}
which is the same as \cref{eq.multiple.source.RIME}. To account for the faceting explicitly, we can write:
\begin{equation}
\Vis_{pq} = \Gjones_p \sum_{\Omega}\bigg( \sum_{s\in\Omega} \Ejones_{sp} \Kjones_{sp} \Bmatrix_{s} \Kjones_{sq}^H \Ejones_{sq}^H \bigg) \Gjones_q^H
\end{equation}
where $\Omega$ indices denote different facets, and $s$ indices denote individual sources. Each source in a given facet is assumed to have similar direction-dependent gains, and their cumulative signal-to-noise allows for better estimations of this direction-dependent gain than would be possible if solving for each source individually. Thus:
\begin{equation}
\Vis_{pq} = \Gjones_p \sum_{\Omega}\bigg( \sum_{s\in\Omega} \Ejones_{\Omega p} \Kjones_{sp} \Bmatrix_{s} \Kjones_{sq}^H \Ejones_{\Omega q}^H \bigg) \Gjones_q^H
\end{equation}

\pg
We begin with direction-independent calibration to find an estimate of $\Gjones_p\forall p$. This first step is done because this part of the Jones chain is the same throughout the sky, and its inverse solution can therefore be applied directly to the visibilities as follows:
\begin{align}
\Vis_{pq}^\mathrm{corr} &= \Ghatjones_p^{-1} \Vis_{pq} \left( \Ghatjones_q^H \right)^{-1}\\
						&= \Ghatjones_p^{-1}\Gjones_p \sum_{\Omega}\bigg( \sum_{s\in\Omega} \Ejones_{\Omega p} \Kjones_{sp} \Bmatrix_{s} \Kjones_{\Omega q}^H \Ejones_{sq}^H \bigg) \Gjones_q^H \left( \Ghatjones_q^H \right)^{-1}
\end{align}
In so doing, we assume that $\left(\Ejones_{sp}\forall s,p\right) =1$ when solving for $\Ghatjones$. In general, this is of course invalid. For the calibrator source, however, we have $Ejones_{sp} =1 \forall p$ by construction. Because we can choose the calibrator source, the presence of a sufficiently bright source (in our case, an unresolved 3C295) will give extremely good estimations of $\Gjones$. We then have $\Ghatjones_p^{-1}\Gjones_p\sim 1$, and
\begin{align}
\Vis_{pq}^\mathrm{corr} &\sim \sum_{\Omega}\left( \sum_{s\in\Omega} \Ejones_{\Omega p} \Kjones_{sp} \Bmatrix_{s} \Kjones_{sq}^H \Ejones_{\Omega q}^H \right)
\end{align}
and direction-dependent calibration then consists of solving for $\Ehatjones_{\Omega p} = \Ghatjones_p^{-1}\Gjones_p\Ejones_{\Omega p} \sim \Ejones_{\Omega p}$ for all facets.

@@

\pg
Direction-dependent calibration was required, as the field of view is large enough that differential ionospheric errors are introduced as a function of direction: the signal from different sources cross different sections of the ionosphere, and the actual gains for these sources are thus different from the gains for the calibrator. They must thus be solved for individually. Assuming that the scale of ionospheric change is of the order of a degree (which is what is observed empirically), then a facet-based approach will give a ``good enough" result for most of the field. The direction-dependent calibration consisted of a physics-based approach, solving for XX and YY correlations along with a rotation matrix simulating differential Faraday rotation. These gain solutions were then smoothed in time and frequency to provide a time-independent, frequency-dependent amplitude solution for each station. The full LOFAR bandwidth is large enough to allow for Clock/TEC separation, which consists of explicitly modeling the two direction-independent phenomena which dominate in the gain's spectral structure. The clock difference between stations\footnote{Each station has a relatively-accurate clock associated with it, save for the Superterp stations which all share a single clock. Like all clocks, they have finite accuracy, and desynchronisation errors thus creep in over time.} introduces a phase error proportional to $\nu$, while differences in Total Electron Content (TEC) along the line of sight of different stations introduce a phase error inversely proportional to $\nu$. With sufficient spectral coverage, these two effects can therefore be separated and modeled individually, thereby improving the phase solution conditioning.

@@

\pg
The direction-dependent calibration was done using 121 facets, spread as shown in \cref{fig.EGS.widefield.regions}. Some facets converge faster or better than others, and so artefacts due to direction-dependent effects may remain in certain areas of the image but not others. Calibration is done solving for full, complex Jones matrices, rather than a physics-based or diagonal-and-rotation matrix.
The final image was imaged with a $uv$-cut, keeping only baselines longer than 100m. This is because shorter baselines were not used during calibration, and so discarded for imaging. The image resolution is $1.5''$, and the image was created using both quality-based weighting schemes and Briggs weighting, with a Briggs factor of $-0.5$. This means that the imaging was skewed slightly in favour of uniform weighting (weighting visibilities by local $uv$-density) rather than natural weighting (same weight for all visibilities). Imaging and calibration were done using the LOFAR beam model used in AWImager \citepads{awimager}, so the image gives integrated flux values rather than apparent flux values. Deconvolution was performed on the full bandwidth\footnote{This improves the conditioning of the imaging inverse problem, as it means that bright structure present across the full band (e.g. hotspots) can be deconvolved with the smaller PSF peak of the hightest-frequency band.} using SSD-clean with an external mask.

\begin{figure}[h!]
\centering
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{image_dirin_SSD_m_c_di_m.int.restored.fits}.png}}
\caption{\label{image.egs.widefield.noddsc} LOFAR wide-field image cutout, calibrated with the initial sky model and no direction-dependent Jones terms.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{image_full_ampphase_di_m.NS.app.restored.fits}.png}}
\caption{\label{image.egs.widefield.doddsc} LOFAR wide-field image cutout, self-calibrated from \cref{image.egs.widefield.noddsc} with direction-dependent Jones terms.}
\end{subfigure}
\caption{\label{image.egs.widefield.ddsc} Wide-field image cutouts of the full primary beam, centred on the EGS, before and after selfcal.}
\end{figure}

@@

\begin{figure}[h!]
\centering
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{widefield.DI}.png}}
\caption{\label{image.egs.widefield.noddsc.zoom} LOFAR wide-field image cutout, calibrated wtih the initial sky model and no direction-dependent Jones terms.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{widefiled.DD}.png}}
\caption{\label{image.egs.widefield.doddsc.zoom} LOFAR wide-field image cutout, self-calibrated from \cref{image.egs.widefield.noddsc} with direction-dependent Jones terms.}
\end{subfigure}
\caption{\label{image.egs.widefield.ddsc.zoom} Cutouts of smaller regions of \cref{image.egs.widefield.ddsc}.}
\end{figure}

\pg
These sources nevertheless remain in the field, as shown in \cref{image.egs.widefield.ddsc.zoom}. Note that the noise level is not constant throughout the image: this is because the array is not uniformly sensitive to the entire sky, but has a specific beam response. Correcting for this beam response in the image-plane gives integrated flux rather than apparent flux. A full catalogue of the sources in this image is given in \cref{appendix.widefield.catalog}, \cref{widefield.survey.catalog} % gives rise to this effect.

\pg
Finally, the quality-based weighting schemes developed in \cref{chapter.paper} were applied throughout this procedure. Noise maps made with and without the weighting scheme are shown in \cref{image.egs.widefield.weights}. A few things are important to note in this comparison. Firstly, the weighting schemes decrease the noise by a factor of $\sim10$: for the noise-map to be visible without weighting schemes, the contrast of the weighted image needed to be pushed way low. For reference, \cref{image.egs.widefield.weighted} is the same image (albeit without the facets shown explicitly) as \cref{fig.EGS.widefield.regions}. The noise distribution is more clearly visible in that figure. Relatedly, by comparing that image and \cref{image.egs.widefield.noweights}, we note that the unweighted image has a different noise per facet, unlike the weighted image. This is due to the fact that a single visibility will be calibrated better or worse depending on the facet which the flux model is used for calibration. It is no coincidence that the facet containing 3C295 is the only one with a noise even remotely comparable to the weighted image's: it is the facet with the best signal-to-noise. This becomes significant when taking into account the fact that the last 2 hours of this observation are much noisier than the first 6: the quality-based weights will decrease their overall contribution \textit{throughout the image}, whereas not using them will lead to different facets having stronger or lesser contributions from these noisy visibilities - and therefore different noise levels. This is also why fringes can clearly be seen in some facets: a single visibility (or a set of a few visibilities) are particularly badly-calibrated, and the associated fringe dominates in the facet. Linking back to \cref{chapter.paper}, because we have a gain solution per visibility \textit{and per facet}, we have one noise-PSF per facet; this suggests that the quality-based weighting scheme could likely be extended to a direction-dependent formalism.

@@

% % % % % % % % % % % % % % % % % % % % % % %
\section{Overlays \& Images}\label{sec.lowresEGS.overlays}

%\subsection{Selected Sources}

\pg
We select 12 sources in the wide-field image, which is shown in \cref{plot.EGS.lofar.widefield}. Our selection is primarily based on whether sources have interesting or peculiar diffuse emission, though some are chosen for being particularly classic examples of radio galaxies. We scan the image from south to north and east to west (scanning upwards then shifting to the right in the image). The positions of the brightest sources ($>1$Jy)are given in \cref{table.egs.sources}.

@@

%\subsection{Images \& Cross-matching}

@@

\begin{figure}[h!]
\centering
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs1.LOFARegs}.png}}
\caption{\label{fig.egs1.lofarim} Cutout of our LOFAR wide-field image centred on EGS-1.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs1.lofar}.png}}
\caption{\label{fig.egs1.sdss.lofoverlay} SDSS image of EGS-1, with \cref{fig.egs1.lofarim} as an overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs1.nvss}.png}}
\caption{\label{fig.egs1.sdss.nvssoverlay} Same image as \cref{fig.egs1.sdss.lofoverlay}, with NVSS overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs1.Wise3.4}.png}}
\caption{\label{fig.egs1.sdss.wiseoverlay} Same image as \cref{fig.egs1.sdss.lofoverlay}, with WISE overlay.}
\end{subfigure}
\caption{\label{fig.egs1} Images \& Overlays for EGS-1.}
\end{figure}

\pg
EGS-1 was selected from the wide-field LOFAR image because it showed interesting diffuse structure. It does not seem to match up to any optical source, nor does it match an infrared source. EGS-1 therefore seems to be best identified as a low-frequency counterpart to NVSS J143706+533225. Its shape seems to include two jets, one much brighter than the other, but both visible in \cref{fig.egs1.lofarim}. This is congruent with a radio galaxy with one Doppler-boosted jet. The absence of an optical counterpart could be due to dust extinction or to insufficient sensitivity.

@@

\begin{figure}[h!]
\centering
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs2.LOFARegs}.png}}
\caption{\label{fig.egs2.lofarim} Cutout of our LOFAR wide-field image centred on EGS-2.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs2.lofar}.png}}
\caption{\label{fig.egs2.sdss.lofoverlay} SDSS image of EGS-2, with \cref{fig.egs2.lofarim} as an overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs2.nvss}.png}}
\caption{\label{fig.egs2.sdss.nvssoverlay} Same image as \cref{fig.egs2.sdss.lofoverlay}, with NVSS overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs2.Wise3.4}.png}}
\caption{\label{fig.egs2.sdss.wiseoverlay} Same image as \cref{fig.egs2.sdss.lofoverlay}, with WISE overlay.}
\end{subfigure}
\caption{\label{fig.egs2} Images \& Overlays for EGS-2.}
\end{figure}

@@

\begin{figure}[h!]
\centering
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs3.LOFARegs}.png}}
\caption{\label{fig.egs3.lofarim} Cutout of our LOFAR wide-field image centred on EGS-3.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs3.lofar}.png}}
\caption{\label{fig.egs3.sdss.lofoverlay} SDSS image of EGS-3, with \cref{fig.egs3.lofarim} as an overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs3.nvss}.png}}
\caption{\label{fig.egs3.sdss.nvssoverlay} Same image as \cref{fig.egs3.sdss.lofoverlay}, with NVSS overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs3.Wise3.4}.png}}
\caption{\label{fig.egs3.sdss.wiseoverlay} Same image as \cref{fig.egs3.sdss.lofoverlay}, with WISE overlay.}
\end{subfigure}
\caption{\label{fig.egs3} Images \& Overlays for EGS-3.}
\end{figure}

@@

\begin{figure}[h!]
\centering
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs4.LOFARegs}.png}}
\caption{\label{fig.egs4.lofarim} Cutout of our LOFAR wide-field image centred on EGS-4.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs4.lofar}.png}}
\caption{\label{fig.egs4.sdss.lofoverlay} SDSS image of EGS-4, with \cref{fig.egs4.lofarim} as an overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs4.nvss}.png}}
\caption{\label{fig.egs4.sdss.nvssoverlay} Same image as \cref{fig.egs4.sdss.lofoverlay}, with NVSS overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs4.Wise3.4}.png}}
\caption{\label{fig.egs4.sdss.wiseoverlay} Same image as \cref{fig.egs4.sdss.lofoverlay}, with WISE overlay.}
\end{subfigure}
\caption{\label{fig.egs4} Images \& Overlays for EGS-4.}
\end{figure}

>>>>

\chapter{Imaging the full LOFAR field of view}\label{section.EGS.lowres}
\minitoc
% % % % % % % % % % % % % % % % % % % % % % %
\section{Aims \& Methodology}

@@

 % % % % % % % % % % % % % % % % % % % % % %
\section{Data Reduction}

\subsection{Data \& Observation Properties}
\pg
The dataset used for this PhD project is part of the LOFAR Surveys KSP, and consists of a number of 8-hour pointings covering as much of the sky visibile to LOFAR as possible. We analyse one of these pointings, an observation performed on the 28th of August 2014 and centred on the Extended Groth Strip. It is an 8-hour-long observation; there exist another 4 8-hour pointings in the same direction, but these data are not reduced in this project. We limit ourselves to analysing only the HBA observation, meaning that we use 365 sub-bands which sample a total bandwidth of 110-182 MHz. We use all Core and Remote LOFAR stations, as well as some of the International stations which were online at the time (specifically the German stations 1-5 and 7, along with the Swedish and the British station).

@@

\pg
The data was acquired through the LOFAR Long-Term Archive, and is thus pre-processed and flagged for RFI using the standard tools \citepads[NDPPP and AOFlagger; see][ respectively]{2018ascl.soft04003V,2010ascl.soft10017O}. The next step is therefore to calibrate the data.

@@

\pg
What's more, Jones matrices can vary over short durations or show peculiar spectral behaviour. Even if interferometric data were averaged massively using baseline-dependent window functions, information on this underlying gain behaviour would be lost - this means that averaging data to reduce its size (and associated reduction wall time) is a risky proposition.

@@

\pg
The calibration strategy consists of a first round of direction-independent calibration followed by direction-dependent self-calibration. In effect, we use the following RIME:
\begin{equation}
\Vis_{pq} = \Gjones_p \bigg( \sum_s \Ejones_{sp} \Kjones_{sp} \Bmatrix_{s} \Kjones_{sq}^H \Ejones_{sq}^H \bigg) \Gjones_q^H
\end{equation}
which is the same as \cref{eq.multiple.source.RIME}. To account for the faceting explicitly, we can write:
\begin{equation}
\Vis_{pq} = \Gjones_p \sum_{\Omega}\left( \Ejones_{\Omega p} \left( \sum_{s\in\Omega}\Kjones_{sp} \Bmatrix_{s} \Kjones_{sq}^H\right) \Ejones_{\Omega q}^H \right) \Gjones_q^H
\end{equation}
where $\Omega$ indices denote different facets, and $s$ indices denote individual sources. Because we make the approximation that the $\Ejones$ term is constant within a facet, we directly write $\Ejones_{\Omega p}$ rather than $\Ejones_{sp}$. Each source in a given facet is assumed to have similar direction-dependent gains, and their cumulative signal-to-noise allows for better estimations of this direction-dependent gain than would be possible if solving for each source individually. 
Thus:
\begin{equation}
\Vis_{pq} = \Gjones_p \sum_{\Omega}\bigg( \sum_{s\in\Omega} \Ejones_{\Omega p} \Kjones_{sp} \Bmatrix_{s} \Kjones_{sq}^H \Ejones_{\Omega q}^H \bigg) \Gjones_q^H
\end{equation}

\pg
We begin with direction-independent calibration to find an estimate of $\Gjones_p\forall p$. This first step is done because this part of the Jones chain is the same throughout the sky, and its inverse solution can therefore be applied directly to the visibilities as follows:
\begin{align}
\Vis_{pq}^\mathrm{corr} &= \Ghatjones_p^{-1} \Vis_{pq} \left( \Ghatjones_q^H \right)^{-1}\\
						&= \Ghatjones_p^{-1}\Gjones_p \sum_{\Omega}\bigg( \sum_{s\in\Omega} \Ejones_{\Omega p} \Kjones_{sp} \Bmatrix_{s} \Kjones_{\Omega q}^H \Ejones_{sq}^H \bigg) \Gjones_q^H \left( \Ghatjones_q^H \right)^{-1}
\end{align}
In so doing, we assume that $\left(\Ejones_{sp}\forall s,p\right) =1$ when solving for $\Ghatjones$. In general, this is of course invalid. For the calibrator source, however, we have $\Ejones_{sp} =1 \forall p$ by construction. Because we can choose the calibrator source, the presence of a sufficiently bright source (in our case, an unresolved 3C295) will give extremely good estimations of $\Gjones$. We then have $\Ghatjones_p^{-1}\Gjones_p\sim \I$, and
\begin{align}
\Vis_{pq}^\mathrm{corr} &\sim \sum_{\Omega}\left( \sum_{s\in\Omega} \Ejones_{\Omega p} \Kjones_{sp} \Bmatrix_{s} \Kjones_{sq}^H \Ejones_{\Omega q}^H \right)\label{eq.vcorr}
\end{align}
and direction-dependent calibration then consists of solving for $\Ehatjones_{\Omega p} = \Ghatjones_p^{-1}\Gjones_p\Ejones_{\Omega p} \sim \Ejones_{\Omega p}$ for all facets -- this means that, if there are errors in the direction-independent gain solutions, they will bias all direction-dependent gain solutions.

@@

\pg
Direction-dependent calibration is required, as the field of view is large enough that differential ionospheric errors are introduced as a function of direction: the signal from different sources cross different sections of the ionosphere, and the actual gains for these sources are thus different from the gains for the calibrator (see \cref{eq.vcorr}). They must thus be solved for individually. Assuming that the scale of ionospheric change is of the order of a degree (which is what is observed empirically), then a facet-based approach will give a ``good enough" result for most of the field. The direction-dependent calibration consisted of a physics-based approach, solving for XX and YY correlations along with a rotation matrix simulating differential Faraday rotation. These gain solutions were then smoothed in time and frequency to provide a time-independent, frequency-dependent amplitude solution for each station. The full LOFAR bandwidth is large enough to allow for Clock/TEC separation, which consists of explicitly modeling the two direction-independent phenomena which dominate in the gain's spectral structure. The clock difference between stations\footnote{Each station has a relatively-accurate clock associated with it, save for the Superterp stations which all share a single clock. Like all clocks, they have finite accuracy, and desynchronisation errors thus creep in over time.} introduces a phase error proportional to $\nu$, while differences in Total Electron Content (TEC) along the line of sight of different stations introduce a phase error inversely proportional to $\nu$. With sufficient spectral coverage, these two effects can therefore be separated and modeled individually, thereby improving the phase solution conditioning.

@@

\pg
The direction-dependent calibration was done using 121 facets, spread as shown in \cref{fig.EGS.widefield.regions}. Some facets converge faster or better than others, and so artefacts due to direction-dependent effects may remain in certain areas of the image but not others. Calibration is done solving for scalar, complex Jones matrices, rather than a physics-based or diagonal-and-rotation matrix.
The final image was imaged with a $uv$-cut, keeping only baselines longer than 100m. This is because shorter baselines were not used during calibration, and so discarded for imaging. The image resolution is $1.5''$, and the image was created using both quality-based weighting schemes and Briggs weighting, with a Briggs factor of $-0.5$. This means that the imaging was skewed slightly in favour of uniform weighting (weighting visibilities by local $uv$-density) rather than natural weighting (same weight for all visibilities). Imaging and calibration were done using the LOFAR beam model used in AWImager \citepads{awimager}, so the image gives integrated flux values rather than apparent flux values. Deconvolution was performed on the full bandwidth\footnote{This improves the conditioning of the imaging inverse problem, as it means that bright structure present across the full band (e.g. hotspots) can be deconvolved with the smaller PSF peak of the hightest-frequency band.} using SSD-clean with an external mask.

\begin{figure}[h!]
\centering
\begin{subfigure}{.8\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{image_dirin_SSD_m_c_di_m.int.restored.fits}.png}}
\caption{\label{image.egs.widefield.noddsc} LOFAR wide-field image cutout, calibrated with the initial sky model and no direction-dependent Jones terms.}
\end{subfigure}
\hfill
\begin{subfigure}{.8\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{image_full_ampphase_di_m.NS.app.restored.fits}.png}}
\caption{\label{image.egs.widefield.doddsc} LOFAR wide-field image cutout, self-calibrated from \cref{image.egs.widefield.noddsc} with direction-dependent Jones terms.}
\end{subfigure}
\caption{\label{image.egs.widefield.ddsc} Wide-field image cutouts of the full primary beam, centred on the EGS, before and after selfcal.}
\end{figure}

@@

\begin{figure}[h!]
\centering
\begin{subfigure}{.8\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{widefield.DI}.png}}
\caption{\label{image.egs.widefield.noddsc.zoom} LOFAR wide-field image cutout, calibrated wtih the initial sky model and no direction-dependent Jones terms.}
\end{subfigure}
\hfill
\begin{subfigure}{.8\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{widefiled.DD}.png}}
\caption{\label{image.egs.widefield.doddsc.zoom} LOFAR wide-field image cutout, self-calibrated from \cref{image.egs.widefield.noddsc} with direction-dependent Jones terms.}
\end{subfigure}
\caption{\label{image.egs.widefield.ddsc.zoom} Cutouts of smaller regions of \cref{image.egs.widefield.ddsc}.}
\end{figure}

\pg
These sources nevertheless remain in the field, as shown in \cref{image.egs.widefield.ddsc.zoom}. Note that the noise level is not constant throughout the image: this is because the array is not uniformly sensitive to the entire sky, but has a specific beam response. Correcting for this beam response in the image-plane gives integrated flux rather than apparent flux. %A full catalogue of the sources in this image is given in \cref{appendix.widefield.catalog}, \cref{widefield.survey.catalog} % gives rise to this effect.

\section{Effect of Quality-based Weighting Scheme}

\pg
Finally, the quality-based weighting scheme developed in \cref{chapter.paper} was applied throughout this procedure. Noise maps made with and without the weighting scheme are shown in \cref{image.egs.widefield.weights}. A few things are important to note in this comparison. Firstly, the weighting schemes decrease the noise by a factor of $\sim10$: for the noise-map to be visible without weighting schemes, the contrast of the weighted image needed to be pushed way low. For reference, \cref{image.egs.widefield.weighted} is the same image (albeit without the facets shown explicitly) as \cref{fig.EGS.widefield.regions}. The noise distribution is more clearly visible in that figure. Relatedly, by comparing that image and \cref{image.egs.widefield.noweights}, we note that the unweighted image has a different noise per facet, unlike the weighted image. This is due to the fact that a single visibility will be calibrated better or worse depending on the facet which the flux model is used for calibration. It is no coincidence that the facet containing 3C295 is the only one with a noise even remotely comparable to the weighted image's: it is the facet with the best signal-to-noise. This becomes significant when taking into account the fact that the last 2 hours of this observation are much noisier than the first 6: the quality-based weights will decrease their overall contribution \textit{throughout the image}, whereas not using them will lead to different facets having stronger or lesser contributions from these noisy visibilities - and therefore different noise levels. This is also why fringes can clearly be seen in some facets: a single visibility (or a set of a few visibilities) are particularly badly-calibrated, and the associated fringe dominates in the facet. Linking back to \cref{chapter.paper}, because we have a gain solution per visibility \textit{and per facet}, we have one noise-PSF per facet; this suggests that the quality-based weighting scheme could likely be extended to a direction-dependent formalism.

@@

% % % % % % % % % % % % % % % % % % % % % % %
\section{Zoo of Radio Sources}\label{sec.lowresEGS.overlays}

\subsection{Selected Sources}

\pg
We select 12 sources in the wide-field image, which is shown in \cref{plot.EGS.lofar.widefield}. Our selection is primarily based on whether sources have interesting or peculiar diffuse emission, though some are chosen for being particularly classic examples of radio galaxies. We scan the image from south to north and east to west (scanning upwards then shifting to the right in the image). The positions of the brightest sources ($>1$Jy)are given in \cref{table.egs.sources}. We insist that this is a very crude scientific analysis, aimed as a showcase for the reader of this manuscript.

@@

\subsection{Images \& Cross-matching}

@@

\begin{figure}[h!]
\centering
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs1.LOFARegs}.png}}
\caption{\label{fig.egs1.lofarim} Cutout of our LOFAR wide-field image centred on EGS-1.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs1.lofar}.png}}
\caption{\label{fig.egs1.sdss.lofoverlay} SDSS image of EGS-1, with \cref{fig.egs1.lofarim} as an overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs1.nvss}.png}}
\caption{\label{fig.egs1.sdss.nvssoverlay} Same image as \cref{fig.egs1.sdss.lofoverlay}, with NVSS overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs1.Wise3.4}.png}}
\caption{\label{fig.egs1.sdss.wiseoverlay} Same image as \cref{fig.egs1.sdss.lofoverlay}, with WISE overlay.}
\end{subfigure}
\caption{\label{fig.egs1} Images \& Overlays for EGS-1. The overlays start at the $2\sigma$ level and end on the image emission peak.}
\end{figure}

\pg
EGS-1 was selected from the wide-field LOFAR image because it showed interesting diffuse structure. It does not seem to match up to any optical source, nor does it match an infrared source. EGS-1 therefore seems to be best identified as a low-frequency counterpart to NVSS J143706+533225. Its shape seems to include two jets, one much brighter than the other, but both visible in \cref{fig.egs1.lofarim}. This could include a radio galaxy with one Doppler-boosted jet, or asymmetry in its environment density. % The absence of an optical counterpart could be due to dust extinction or to insufficient sensitivity.

@@

\begin{figure}[h!]
\centering
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs2.LOFARegs}.png}}
\caption{\label{fig.egs2.lofarim} Cutout of our LOFAR wide-field image centred on EGS-2.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs2.lofar}.png}}
\caption{\label{fig.egs2.sdss.lofoverlay} SDSS image of EGS-2, with \cref{fig.egs2.lofarim} as an overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs2.nvss}.png}}
\caption{\label{fig.egs2.sdss.nvssoverlay} Same image as \cref{fig.egs2.sdss.lofoverlay}, with NVSS overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs2.Wise3.4}.png}}
\caption{\label{fig.egs2.sdss.wiseoverlay} Same image as \cref{fig.egs2.sdss.lofoverlay}, with WISE overlay.}
\end{subfigure}
\caption{\label{fig.egs2} Images \& Overlays for EGS-2. The overlays start at the $5\sigma$ level and end on the image emission peak.}
\end{figure}

@@

\begin{figure}[h!]
\centering
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs3.LOFARegs}.png}}
\caption{\label{fig.egs3.lofarim} Cutout of our LOFAR wide-field image centred on EGS-3.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs3.lofar}.png}}
\caption{\label{fig.egs3.sdss.lofoverlay} SDSS image of EGS-3, with \cref{fig.egs3.lofarim} as an overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs3.nvss}.png}}
\caption{\label{fig.egs3.sdss.nvssoverlay} Same image as \cref{fig.egs3.sdss.lofoverlay}, with NVSS overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs3.Wise3.4}.png}}
\caption{\label{fig.egs3.sdss.wiseoverlay} Same image as \cref{fig.egs3.sdss.lofoverlay}, with WISE overlay.}
\end{subfigure}
\caption{\label{fig.egs3} Images \& Overlays for EGS-3. The overlays start at the $1.5\sigma$ level and end on the image emission peak.}
\end{figure}

@@

\begin{figure}[h!]
\centering
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs4.LOFARegs}.png}}
\caption{\label{fig.egs4.lofarim} Cutout of our LOFAR wide-field image centred on EGS-4.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs4.lofar}.png}}
\caption{\label{fig.egs4.sdss.lofoverlay} SDSS image of EGS-4, with \cref{fig.egs4.lofarim} as an overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs4.nvss}.png}}
\caption{\label{fig.egs4.sdss.nvssoverlay} Same image as \cref{fig.egs4.sdss.lofoverlay}, with NVSS overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs4.Wise3.4}.png}}
\caption{\label{fig.egs4.sdss.wiseoverlay} Same image as \cref{fig.egs4.sdss.lofoverlay}, with WISE overlay.}
\end{subfigure}
\caption{\label{fig.egs4} Images \& Overlays for EGS-4. The overlays start at the $2\sigma$ level and end on the image emission peak.}
\end{figure}
